---
title: "DSC 200 Lab Project -Term 2232"
date: "`r Sys.Date()`"
output: github_document
---

**Student Name:< Aminah Mohammead Almutairi>**

**Student ID:<2221001947>**


**Deadline:** 23:59 on Sunday, 19 May 2024

**Total Points:** 20


## Loading Packages

```{r load-packages, message = FALSE}

library(tidyverse)
library(openintro)
library(ggrepel)
```


## Tasks

`1. (2 points)
```{r}
# Count the number of pets
num_pets <- nrow(seattlepets)
num_pets
```
Write your narrative here
nrow() is a function in R that returns the number of rows in a dataset.
seattlepets is assumed to be the name of your dataset containing information about pets.
The result of nrow(seattlepets) is assigned to the variable num_pets, which will contain the total number of pets in the dataset.



`2. (2 points)

```{r}
# Count the number of variables
num_variables <- ncol(seattlepets)
num_variables
```
Write your narrative below
ncol() is a function in R that returns the number of columns in a dataset.
seattlepets is assumed to be the name of your dataset containing information about pets.
The result of ncol(seattlepets) is assigned to the variable num_variables, which will contain the total number of variables (columns) in the dataset.


`3. (2 points)
```{r}
# Count the frequency of each species
species_count <- table(seattlepets$species)
species_count
```
Write your narrative here
table() is a function in R used to tabulate the counts of data values.
seattlepets$species assumes that "seattlepets" is the name of your dataset and "species" is the column containing species information.
table(seattlepets$species) creates a table that shows the frequency of each unique species in the dataset.
The result is stored in the variable species_count, which holds the frequency counts of each species.

`4. (2 points)
```{r}
# Count the frequency of each pet name
pet_name_frequency <- table(seattlepets$animal_name)

# Sort the frequency table in descending order
sorted_pet_names <- sort(pet_name_frequency, decreasing = TRUE)

# Display the top ten most common pet names
head(sorted_pet_names,10)
```
Write your narrative here
table(seattlepets$animal_name) creates a frequency table of pet names in the "seattlepets" dataset.
sort() sorts the frequency table in descending order of counts. By default, it sorts in ascending order, but by specifying decreasing = TRUE, it sorts in descending order.
head(sorted_pet_names, 10) displays the first ten entries of the sorted frequency table, which represent the top ten most common pet names.

`5. (2 points)
```{r}
# Filter records for species Pig
pig_records <- filter(seattlepets, species == "Pig")

# Arrange the filtered records by pet names (animal_name)
sorted_pig_records <- arrange(pig_records, animal_name)

# Display the sorted pig records
sorted_pig_records
```
Write your narrative here

filter(seattlepets, species == "Pig") filters the dataset "seattlepets" to include only records where the species is labeled as "Pig". This creates a new dataset named "pig_records".
arrange(pig_records, animal_name) sorts the "pig_records" dataset by the column "animal_name" in ascending order by default.
The sorted pig records are then stored in the variable sorted_pig_records.
sorted_pig_records is then displayed, showing the filtered and sorted records of pigs.


`6. (2 points)
```{r}
# Select only pet name (animal_name) and primary breed for species Goat
goat_records <- select(filter(seattlepets, species == "Goat"), animal_name, primary_breed)

# Arrange the selected records by pet names (animal_name)
sorted_goat_records <- arrange(goat_records, animal_name)

# Display the sorted goat records
sorted_goat_records
```
Write your narrative here

filter(seattlepets, species == "Goat") filters the dataset "seattlepets" to include only records where the species is labeled as "Goat".
select() is used to choose specific columns from the filtered dataset. In this case, it selects the columns "animal_name" (pet name) and "primary_breed".
The selected records are stored in the variable goat_records.
arrange(goat_records, animal_name) sorts the "goat_records" dataset by the column "animal_name" in ascending order by default.
The sorted goat records are stored in the variable sorted_goat_records.
Finally, sorted_goat_records is displayed, showing the selected and sorted records for goats.



`7. (2 points)
```{r}
library(dplyr)

# Concatenate animal_name and species into a single column named pet
seattlepets <- mutate(seattlepets, pet = paste(animal_name, species, sep = " - "))

# Select license_number and pet, then arrange by pet
sorted_records <- select(seattlepets, license_number, pet) %>%
                  arrange(pet)

# Display the sorted records
sorted_records
```
Write your narrative here

The mutate() function from the dplyr package is used to create a new column named "pet" in the "seattlepets" dataset. This column is created by concatenating the columns "animal_name" and "species" with a separator " - ".
select(seattlepets, license_number, pet) selects the columns "license_number" and "pet" from the modified dataset.
The %>% operator (pipe operator) is used to pass the result of the previous operation (select()) as the first argument to the next operation (arrange()).
arrange(pet) sorts the dataset by the "pet" column in ascending order.
The sorted records are stored in the variable sorted_records.
Finally, sorted_records is displayed, showing the selected and sorted records.

`8. (2 points)
```{r}
library(ggplot2)

# Create a bar plot of species counts
species_plot <- ggplot(seattlepets, aes(x = species)) +
                geom_bar() +
                labs(title = "Counts of Species", x = "Species", y = "Count") +
                theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Display the plot
print(species_plot)
```
Write your narrative here
ggplot(seattlepets, aes(x = species)) initializes a ggplot object with the dataset "seattlepets" and aesthetic mapping specifying that the x-axis represents the "species" column.
geom_bar() adds a layer of bars to the plot, where each bar represents the count of occurrences of a particular species.
labs() sets the title and labels for the axes.
In the theme() function, axis.text.x = element_text(angle = 45, hjust = 1) adjusts the appearance of the x-axis labels by rotating them 45 degrees clockwise and adjusting their horizontal justification to 1 (right-aligned).
The variable species_plot holds the ggplot object. Finally, print(species_plot) displays the plot.



`9. (2 points)
```{r}
top_10_names <- seattlepets %>% 
filter(animal_name %in% c( "Lucy"  , "Charlie" , "Luna" , "Bella" , "Max"    , 
                           "Daisy" , "Molly"   , "Jack" , "Lily"  , "Stella"))
top_10_names
```
`a.	What does the above code chunk do?
```{r}
# Create a subset of data for the top 10 most common pet names
top_10_names <- c("Lucy", "Charlie", "Luna", "Bella", "Max", "Cooper", "Daisy", "Buddy", "Molly", "Stella")
top_10_subset <- filter(seattlepets, animal_name %in% top_10_names)
```
`b. Plot the counts of the pet names (animal_name) in top_10_names
```{r}
# Plot the counts of the pet names segmented by species
pet_names_plot <- ggplot(top_10_subset, aes(x = animal_name, fill = species)) +
                  geom_bar(position = "dodge") +
                  labs(title = "Counts of Top 10 Pet Names Segmented by Species", x = "Pet Name", y = "Count") +
                  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Display the plot
print(pet_names_plot)
```
`10. (2 points)
```{r eval = TRUE, echo = FALSE, fig.fullwidth = TRUE}
# data prep
name_props <- seattlepets %>% 
  filter(
    !is.na(animal_name),
    species %in% c("Dog", "Cat")
    ) %>%
  group_by(species) %>% 
  count(animal_name, sort = TRUE) %>%
  mutate(prop = n / sum(n))
    
cat_name_props <- name_props %>%
  filter(species == "Cat") %>%
  rename(cat_prop = prop) %>%
  slice(1:30)

dog_name_props <- name_props %>%
  filter(species == "Dog") %>%
  rename(dog_prop = prop) %>%
  slice(1:30)
    
comb_name_props <- inner_join(cat_name_props, dog_name_props, 
                              by = "animal_name") %>%
  ungroup() %>%
  select(animal_name, cat_prop, dog_prop)
    
# create viz
ggplot(comb_name_props, aes(x = cat_prop, y = dog_prop)) +
  geom_abline(intercept = 0, 
              color = COL["lgray","full"], 
              alpha = 0.8, 
              size = 1.5) +
  geom_text_repel(aes(label = animal_name), 
                  segment.color = COL["gray","full"], 
                  seed = 291252, max.iter = 10000) +
  geom_point(color = COL["blue","full"], alpha = 0.8) +
  theme_minimal() +
  labs(x = "Proportion of cats", y = "Proportion of dogs") +
  xlim(0.002, 0.01) +
  ylim(0.002, 0.01) +
  ggimage::geom_emoji(
      image = "1f436", aes(x = 0.003, y = 0.009), size = 0.1
      ) +
  ggimage::geom_emoji(
      image = "1f431", aes(x = 0.009, y = 0.003), size =0.1)
```
`What names are more common for cats than dogs? The ones above the line or the ones below the line?

`Answer here...Dog:Lucy.Cat:Luna...............



`Is the relationship between the two variables (proportion of cats with a given name and proportion of dogs with a given name) positive or negative? What does this mean in context of the data?

`Answer here .Positive,as the popularity of a name for dogs increases,so does the popularity of that name for cats............
